Subject-specific curves for time series forecasting of smart meter demand

### 2019-11-03 sscts-paper-1.Rmd
subject-specific

To justify our final model that we use for scenario analysis we test several models that can be thought of as simpler versions. Each model has a change introduced and the improvement in performance is used as justification for each. As a starting benchmark, we fit individual linear regression models to each building. The second set of individual models incorporate natural splines to model the temperature and demand relationships. The third, fourth and fifth models are random intercept, random slope and subject specific curve models. Finally, the sixth model is a subject specific curve model that includes building attributes as fixed effects. We also fit a naive model that uses the previous day's observed values as a benchmark. This is summarised in Table \@ref(tab:model-frameworks).

@@

Model                        | Abbreviation | Description              | Weather Variables
-----------------------------|--------------|--------------------------|-----------
Naive                        | Naive | Naive forecasting model | None
Individual linear regression | ILR  | Linear regression models fit to each building | Current temperature
Individual natural splines   | INS  | Natural spline models fit to each building | Feature selection
Pooled regression            | PR | Regression model. Used for feature selection | All/Feature selection
Random intercept             | RI   | Mixed effects model with random intercept | Feature selection
Random intercept and slope   | RI   | Mixed effects model with random intercept and slope | Feature selection
Subject specific curves      | SSC  | Mixed effects model with subject specific curves | Feature selection
Subject specific curves with attributes | SSCATTR | SSC with building attributes included as fixed effects | Feature selection

@@

Originally, random effects were incorporated into each of the lagged weather variables, but this resulted in very poor fits presumably due to the high dimensionality. Instead, as with the PR model, we model the population relationship for all selected predictor variables and allow for subject specific differences using random intercepts, random slopes and subject specific curves (based on current temperature).

@@

#### Subject specific curves model

As the relationship between demand and temperature can be quite non-linear we also explore modelling the subject specific differences in the temperature and energy relationship using splines. We call this model the subject specific curve (SSC) model in keeping with @Durban2005-lk. Note that we have modified their model to work with natural cubic splines as this gives a better fit when modelling the temperature and electricity relationship compared to penalized linear splines (Figure \@ref(fig:plot-linear-cubic-splines)). It is given by
$$
\begin{gathered}
\log y_{bt} = \sum_{x_{b,i,t} \in \mathcal{P}_t } f_{i,p}(x_{b,i,t}) + g_{b,p}( w_{b,0,t} ) + \epsilon_{b,t}, \\
f_{i,p}(x) = \sum^K_{k=1} \beta_{i,p,k} (x - \kappa_k)_+^3, \quad g_{b,p}(x) = \sum^K_{k=1} u_{b,p,k} (x - \kappa_k)_+^3, \\
\epsilon_{b,t} \sim N(0, \sigma^2_{\epsilon, p}), \quad
u_{b,p,k} \sim N(0, \sigma^2_{u,p}).
\end{gathered}
$$

@@

#### Subject specific curves with attributes model

As we wish to carry out scenario analysis we introduce several new variables into our model. These variables are the set of building attributes discussed in Section \@ref(attribute-data) which we denote by $\mathcal{A}$. We treat each of these attributes as a fixed effect. We refer to this model as the subject specific curves with attributes (SSCATTR) model.

>>>>

To justify our final model that we use for scenario analysis we test several models that can be thought of as simpler versions. Each model has a change introduced and the improvement in performance is used as justification for each. As a starting benchmark, we fit individual linear regression models to each building. The second set of individual models incorporate natural splines to model the temperature and demand relationships. The third, fourth and fifth models are random intercept, random slope and subject-specific curve models. Finally, the sixth model is a subject-specific curve model that includes building attributes as fixed effects. We also fit a naive model that uses the previous day's observed values as a benchmark. This is summarised in Table \@ref(tab:model-frameworks).

@@

Model                        | Abbreviation | Description              | Weather Variables
-----------------------------|--------------|--------------------------|-----------
Naive                        | Naive | Naive forecasting model | None
Individual linear regression | ILR  | Linear regression models fit to each building | Current temperature
Individual natural splines   | INS  | Natural spline models fit to each building | Feature selection
Pooled regression            | PR | Regression model. Used for feature selection | All/Feature selection
Random intercept             | RI   | Mixed effects model with random intercept | Feature selection
Random intercept and slope   | RI   | Mixed effects model with random intercept and slope | Feature selection
Subject-specific curves      | SSC  | Mixed effects model with subject-specific curves | Feature selection
Subject-specific curves with attributes | SSCATTR | SSC with building attributes included as fixed effects | Feature selection

@@

Originally, random effects were incorporated into each of the lagged weather variables, but this resulted in very poor fits presumably due to the high dimensionality. Instead, as with the PR model, we model the population relationship for all selected predictor variables and allow for subject-specific differences using random intercepts, random slopes and subject-specific curves (based on current temperature).

@@

#### Subject-specific curves model

As the relationship between demand and temperature can be quite non-linear we also explore modelling the subject-specific differences in the temperature and energy relationship using splines. We call this model the subject-specific curve (SSC) model in keeping with @Durban2005-lk. Note that we have modified their model to work with natural cubic splines as this gives a better fit when modelling the temperature and electricity relationship compared to penalized linear splines (Figure \@ref(fig:plot-linear-cubic-splines)). It is given by
$$
\begin{gathered}
\log y_{bt} = \sum_{x_{b,i,t} \in \mathcal{P}_t } f_{i,p}(x_{b,i,t}) + g_{b,p}( w_{b,0,t} ) + \epsilon_{b,t}, \\
f_{i,p}(x) = \sum^K_{k=1} \beta_{i,p,k} (x - \kappa_k)_+^3, \quad g_{b,p}(x) = \sum^K_{k=1} u_{b,p,k} (x - \kappa_k)_+^3, \\
\epsilon_{b,t} \sim N(0, \sigma^2_{\epsilon, p}), \quad
u_{b,p,k} \sim N(0, \sigma^2_{u,p}).
\end{gathered}
$$

@@

#### Subject-specific curves with attributes model

As we wish to carry out scenario analysis we introduce several new variables into our model. These variables are the set of building attributes discussed in Section \@ref(attribute-data) which we denote by $\mathcal{A}$. We treat each of these attributes as a fixed effect. We refer to this model as the subject-specific curves with attributes (SSCATTR) model.
